---
title: "Nighttime Lights Analysis"
author: "Data Lab"
date: "9/20/2024"
format:
  html:
    toc: true
    code-fold: true
knitr:
  opts_chunk:
    warning: false
    message: false
---

```{r}
#| code-fold: true

# PARAMETERS
root_dir     <- "~/Desktop"    # Where to put NTL folder
folder_name  <- "NTL Analysis" # The code will create this folder
iso3_code    <- "SDN"          # ISO3 code for country
pc_base_year <- 2019           # Base year for % change maps
nasa_bearer <- "BEARER-TOKEN"  # NASA Bearer Token
month_start <- "2012-01-01"    # Start month to query 
month_end   <- "2012-01-01"    # End month to query

```

Nighttime lights have become a commonly used resource to estimate changes in local economic activity. This document analyzes nighttime lights within the country.

# Data

We use nighttime lights data from [VIIRS Black Marble](https://blackmarble.gsfc.nasa.gov/). Raw nighttime lights data requires correction due to cloud cover and stray light, such as lunar light. The Black Marble dataset applies [advanced algorithms](https://www.sciencedirect.com/science/article/pii/S003442571830110X) to correct raw nighttime light values and calibrate data so that trends in lights over time can be meaningfully analyzed. From VIIRS Black Marble, we use data from January 2012 through present—where data is available at a 500-meter resolution.

# Methodology

Within different units of analysis (e.g, administrative units) we use the sum of nighttime lights. Where relevant, we distinguish between lights generated from gas flaring and lights removing gas flaring. We use the [World Bank's Global Gas Flaring Tracker](https://www.worldbank.org/en/programs/gasflaringreduction/global-flaring-data) which indicates the location of gas flaring locations. When removing gas flaring lights, we remove lights within 10km of a gas flaring location; when looking at lights in gas flaring locations, we take the sum of lights within 10km of gas flaring locations. 

# Code Setup

```{r}
#| warning: false
#| message: false

#### Packages
library(tidyverse)
library(sf)
library(leaflet)
library(leaflet.providers)
library(ggpubr)
library(terra)
library(tidyterra)
library(gtools)
library(readxl)
library(janitor)
library(geodata)
library(exactextractr)
library(blackmarbler)
library(dplyr)
library(readxl)
library(janitor)
library(ggpubr)
library(WDI)
library(kableExtra)
library(terra)
library(DT)
library(countrycode)

## Iso2 Code
iso2_code <- countrycode(iso3_code, origin = "iso3c", destination = "iso2c")


## Filepath
proj_dir     <- file.path("~/Desktop", folder_name)

data_dir       <- file.path(proj_dir, "data")
ntl_bm_dir     <- file.path(data_dir, "Nighttime Lights BlackMarble")
gadm_dir       <- file.path(data_dir, "GADM")
gas_flare_dir  <- file.path(data_dir, "gas-flaring")

## Create
dir.create(proj_dir)
dir.create(data_dir)
dir.create(ntl_bm_dir)
dir.create(gadm_dir)
dir.create(gas_flare_dir)

#### Bearer
nasa_bearer <- read_csv("~/Dropbox/bearer_bm.csv") %>%
  pull(token)

#### Theme
theme_manual <- theme_classic2() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        plot.title = element_text(face = "bold"))

#### GADM Boundaries
for(i in 0:3){
  for(iso3_i in iso3_code){
    gadm(iso3_code, level=i, path = file.path(gadm_dir))
  }
} 

#### Load GADM Functions
load_gadm <- function(i){
  roi_sf <- file.path(gadm_dir, 'gadm') %>%
    list.files(full.names = T) %>%
    str_subset(paste0("_",i,"_pk.rds")) %>%
    map_df(function(file_path){
      readRDS(file_path) %>%
        st_as_sf()
    })
}

#### Elevation
dir.create(file.path(data_dir, "elevation"))

OUT_PATH <- file.path(data_dir, "elevation", "elev.tiff")

if(!file.exists(OUT_PATH)){
  
  if(length(iso3_code) > 1){
    elev_r <- elevation_30s(country=iso3_code, path=tempdir())
    
  } else{
    elev_list <- lapply(iso3_code, function(iso3_code_i){
      elevation_30s(country=iso3_code_i, path=tempdir())
    }) 
    elev_r <- do.call(mosaic, c(elev_list, fun="mean"))
  }
  
  writeRaster(elev_r, OUT_PATH)
  
}

#### Months to Query
ntl_months <- seq.Date(from = ymd(month_start), 
                       to = ymd(month_end), 
                       by = "month") %>%
  as.character()

#### Colors
wbg_color_light <- "#1AA1DB"
wbg_color_dark <- "#02224B"
```

# Prep Data

## Download Cities

We use a dataset from [Geonames](https://public.opendatasoft.com/explore/dataset/geonames-all-cities-with-a-population-1000/export/?flg=en-us&disjunctive.cou_name_en&sort=name) which indicates the location of cities with over 1,000 people. 

```{r}
#| warning: false
#| message: false

dir.create(file.path(data_dir, "cities"))

## Cities
OUT_FILE_ALL <- file.path(data_dir, "cities", "cities_all.Rds")

if(!file.exists(OUT_FILE_ALL)){
  # https://public.opendatasoft.com/explore/dataset/geonames-all-cities-with-a-population-1000/export/?flg=en-us&disjunctive.cou_name_en&sort=name
  city_df <- read_delim("https://public.opendatasoft.com/api/explore/v2.1/catalog/datasets/geonames-all-cities-with-a-population-1000/exports/csv?lang=en&timezone=America%2FNew_York&use_labels=true&delimiter=%3B", delim = ";")
  
  saveRDS(city_df, OUT_FILE_ALL)
} 

## Country Cities
OUT_FILE_COUNTRY_DF      <- file.path(data_dir, "cities", "cities_country_df.Rds")
OUT_FILE_COUNTRY_SF      <- file.path(data_dir, "cities", "cities_country_sf.Rds")
OUT_FILE_COUNTRY_5K_SF   <- file.path(data_dir, "cities", "cities_country_5km_sf.Rds")
OUT_FILE_COUNTRY_10K_SF  <- file.path(data_dir, "cities", "cities_country_10km_sf.Rds")
OUT_FILE_COUNTRY_BUFF_SF <- file.path(data_dir, "cities", "cities_country_buff_sf.Rds")

if(!file.exists(OUT_FILE_COUNTRY_SF)){
  
  ## Cleanup
  city_country_df <- city_df[city_df$`Country Code` %in% iso2_code,] %>%
    clean_names()
  
  city_country_df <- city_country_df %>%
    separate(coordinates, into = c("lat", "lon"), sep = ",")
  
  ## Spatially Define
  city_country_sf <- city_country_df %>%
    st_as_sf(coords = c("lon", "lat"),
             crs = 4326)
  
  ## Buffer
  city_country_5km_sf <- city_country_sf %>%
    st_buffer(dist = 5000)
  
  city_country_10km_sf <- city_country_sf %>%
    st_buffer(dist = 10000)
  
  ## Dynamic Buffer
  city_country_sf <- city_country_sf %>%
    dplyr::mutate(buffer = case_when(
      population >= 10000000 ~ 30,
      population >= 5000000 ~ 25,
      population >= 1000000 ~ 20,
      population >= 500000 ~ 15,
      population >= 100000 ~ 10,
      TRUE ~ 5
    ))
  
  city_sub_30km_buff_sf <- city_country_sf[city_country_sf$buffer %in% 30,] %>% st_buffer(dist = 30*1000)
  city_sub_25km_buff_sf <- city_country_sf[city_country_sf$buffer %in% 25,] %>% st_buffer(dist = 25*1000)
  city_sub_20km_buff_sf <- city_country_sf[city_country_sf$buffer %in% 20,] %>% st_buffer(dist = 20*1000)
  city_sub_15km_buff_sf <- city_country_sf[city_country_sf$buffer %in% 15,] %>% st_buffer(dist = 15*1000)
  city_sub_10km_buff_sf <- city_country_sf[city_country_sf$buffer %in% 10,] %>% st_buffer(dist = 10*1000)
  city_sub_5km_buff_sf  <- city_country_sf[city_country_sf$buffer %in% 5,]  %>% st_buffer(dist = 5*1000)
  
  city_buff_sf <- bind_rows(city_sub_30km_buff_sf,
                            city_sub_25km_buff_sf,
                            city_sub_20km_buff_sf,
                            city_sub_15km_buff_sf,
                            city_sub_10km_buff_sf,
                            city_sub_5km_buff_sf)
  
  
  ## Export
  saveRDS(city_country_df, OUT_FILE_COUNTRY_DF)
  
  saveRDS(city_country_5km_sf,  OUT_FILE_COUNTRY_5K_SF)
  saveRDS(city_country_10km_sf, OUT_FILE_COUNTRY_10K_SF)
  saveRDS(city_buff_sf,         OUT_FILE_COUNTRY_BUFF_SF)
  saveRDS(city_country_sf,      OUT_FILE_COUNTRY_SF)
  
} 

```


## Download and Append gas flaring

We download and append gas flaring data from the [World Bank Global Gas Flaring Database](https://datacatalog.worldbank.org/search/dataset/0037743).

```{r}
#| warning: false
#| message: false

dir.create(file.path(gas_flare_dir, "rawdata"))
dir.create(file.path(gas_flare_dir, "finaldata"))

#### Download data
# https://datacatalog.worldbank.org/search/dataset/0037743
if(!file.exists(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2017.xlsx"))){
  download.file(url = "https://datacatalogfiles.worldbank.org/ddh-published/0037743/DR0045623/viirs_global_flaring_d.7_slope_0.029353_2017_web_v1.xlsx?versionId=2023-01-18T20:03:32.2273754Z", 
                destfile = file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2017.xlsx"), 
                mode = "wb")
}

if(!file.exists(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2018.xlsx"))){
  download.file(url = "https://datacatalogfiles.worldbank.org/ddh-published/0037743/DR0045622/viirs_global_flaring_d.7_slope_0.029353_2018_web.xlsx?versionId=2023-01-18T20:02:43.3965005Z", 
                destfile = file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2018.xlsx"), 
                mode = "wb")
}

if(!file.exists(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2019.xlsx"))){
  download.file(url = "https://datacatalogfiles.worldbank.org/ddh-published/0037743/DR0045621/viirs_global_flaring_d.7_slope_0.029353_2019_web_v20201114-3.xlsx?versionId=2023-01-18T20:03:09.2456111Z", 
                destfile = file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2019.xlsx"), 
                mode = "wb")
}

if(!file.exists(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2020.xlsx"))){
  download.file(url = "https://datacatalogfiles.worldbank.org/ddh-published/0037743/DR0084248/2020%20Global%20Gas%20Flaring%20Volumes.xlsx?versionId=2023-01-18T20:03:53.8309309Z", 
                destfile = file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2020.xlsx"), 
                mode = "wb")
}

if(!file.exists(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2021.xlsx"))){
  download.file(url = "https://datacatalogfiles.worldbank.org/ddh-published/0037743/DR0087112/2021%20Global%20Gas%20Flaring%20Volumes.xlsx?versionId=2023-01-18T20:02:21.4951166Z", 
                destfile = file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2021.xlsx"), 
                mode = "wb")
}

#### Append data
clean_data <- function(x){
  x %>% 
    clean_names() %>% 
    dplyr::filter(iso_code %in% iso3_code)
} 

df_2021 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2021.xlsx"), 2) %>% clean_data()

df_2020_1 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2020.xlsx"), 1) %>% clean_data()
df_2020_2 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2020.xlsx"), 2) %>% clean_data()
df_2020_3 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2020.xlsx"), 3) %>% clean_data()

df_2019 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2019.xlsx"), 1) %>% clean_data()

df_2018_4 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2018.xlsx"), 4) %>% clean_data()
df_2018_5 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2018.xlsx"), 5) %>% clean_data()
df_2018_6 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2018.xlsx"), 6) %>% clean_data()

df_2017_1 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2017.xlsx"), 1) %>% clean_data()
df_2017_2 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2017.xlsx"), 2) %>% clean_data()
df_2017_3 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2017.xlsx"), 3) %>% clean_data()

gs_df <- bind_rows(
  df_2021,
  df_2020_1,
  df_2020_2,
  df_2020_3,
  df_2019,
  df_2018_4,
  df_2018_5,
  df_2018_6,
  df_2017_1,
  df_2017_2,
  df_2017_3
)

gs_df <- gs_df %>%
  dplyr::select(latitude, longitude) %>%
  distinct() %>%
  dplyr::mutate(uid = 1:n())

gs_sf <- gs_df %>% 
  st_as_sf(coords = c("longitude", "latitude"),
           crs = 4326)

gs_5km_sf <- gs_sf %>%
  st_buffer(dist = 5000)

gs_10km_sf <- gs_sf %>%
  st_buffer(dist = 10000)

saveRDS(gs_df, file.path(gas_flare_dir, "finaldata", "gas_flare_locations.Rds"))
saveRDS(gs_5km_sf, file.path(gas_flare_dir, "finaldata", "gas_flare_locations_5km.Rds"))
saveRDS(gs_10km_sf, file.path(gas_flare_dir, "finaldata", "gas_flare_locations_10km.Rds"))
```


## Download Nighttime Lights

```{r}
#| warning: false
#| message: false

#### NTL
roi_sf <- load_gadm(0)

dir.create(file.path(ntl_bm_dir, "rasters"))
dir.create(file.path(ntl_bm_dir, "rasters", "annual"))
dir.create(file.path(ntl_bm_dir, "rasters", "monthly"))

r <- bm_raster(roi_sf = roi_sf,
               product_id = "VNP46A4",
               date = 2012:2023,
               bearer = nasa_bearer,
               output_location_type = "file",
               file_dir = file.path(ntl_bm_dir, "rasters", "annual"))

r <- bm_raster(roi_sf = roi_sf,
               product_id = "VNP46A3",
               date = ntl_months,
               bearer = nasa_bearer,
               output_location_type = "file",
               file_dir = file.path(ntl_bm_dir, "rasters", "monthly"))

```

## Aggregate Nighttime Lights
```{r}
#| warning: false
#| message: false

# Individual Files -------------------------------------------------------------
dir.create(file.path(ntl_bm_dir, "aggregated_individual_files"))

#### Setup loops
for(unit in c("adm0", 
              "adm1", 
              "adm2",
              "adm3",
              "cities5km",
              "cities10km",
              "citiesbuff")){
  for(time_type in c("annual", "monthly")){
    
    if(time_type == "annual"){
      time_vec <- 2012:2023
    }
    
    if(time_type == "monthly"){
      time_vec <- ntl_months
    }
    
    for(time_i in time_vec){
      
      #### Only process if file doesn't exist
      dir.create(file.path(ntl_bm_dir, "aggregated_individual_files", unit))
      dir.create(file.path(ntl_bm_dir, "aggregated_individual_files", unit, time_type))
      
      OUT_FILE <- file.path(ntl_bm_dir, "aggregated_individual_files", unit, time_type, 
                            paste0("ntl_", time_i, ".Rds"))
      
      if(!file.exists(OUT_FILE)){
        
        #### Load Unit
        if(unit == "adm0"){
          roi_sf <- load_gadm(0)
        }
        
        if(unit == "adm1"){
          roi_sf <- load_gadm(1)
        }
        
        if(unit == "adm2"){
          roi_sf <- load_gadm(2)
        }
        
        if(unit == "adm3"){
          roi_sf <- load_gadm(3)
        }
        
        if(unit == "cities5km"){
          roi_sf <- readRDS(file.path(data_dir, "cities", "cities_country_5km_sf.Rds"))
        }
        
        if(unit == "cities10km"){
          roi_sf <- readRDS(file.path(data_dir, "cities", "cities_country_10km_sf.Rds"))
        }
        
        if(unit == "citiesbuff"){
          roi_sf <- readRDS(file.path(data_dir, "cities", "cities_country_buff_sf.Rds"))
        }
        
        
        #### Load Raster
        if(time_type == "annual"){
          bm_product <- "VNP46A4"
          time_raster_i <- time_i 
        }
        
        if(time_type == "monthly"){
          bm_product <- "VNP46A3"
          
          time_raster_i <- time_i %>% 
            substring(1, 7) %>%
            str_replace_all("-", "_")
        }
        
        ntl_r <- rast(file.path(ntl_bm_dir, "rasters", time_type, 
                                paste0(bm_product,"_NearNadir_Composite_Snow_Free_qflag_t",time_raster_i,".tif")))
        
        #### Gas Flaring Mask
        gs_5km_sf  <- readRDS(file.path(gas_flare_dir, "finaldata", "gas_flare_locations_5km.Rds"))
        gs_10km_sf <- readRDS(file.path(gas_flare_dir, "finaldata", "gas_flare_locations_10km.Rds"))
        
        ntl_gf_5km_r   <- mask(ntl_r, gs_5km_sf)
        ntl_nogf_5km_r <- mask(ntl_r, gs_5km_sf, inverse = T)
        
        ntl_gf_10km_r   <- mask(ntl_r, gs_10km_sf)
        ntl_nogf_10km_r <- mask(ntl_r, gs_10km_sf, inverse = T)
        
        #### Extract NTL
        roi_sf$ntl_sum          <- exact_extract(ntl_r, roi_sf, "sum", progress = F)
        
        roi_sf$ntl_gf_5km_sum   <- exact_extract(ntl_gf_5km_r, roi_sf, "sum", progress = F)
        roi_sf$ntl_nogf_5km_sum <- exact_extract(ntl_nogf_5km_r, roi_sf, "sum", progress = F)
        
        roi_sf$ntl_gf_10km_sum   <- exact_extract(ntl_gf_10km_r, roi_sf, "sum", progress = F)
        roi_sf$ntl_nogf_10km_sum <- exact_extract(ntl_nogf_10km_r, roi_sf, "sum", progress = F)
        
        #### Cleanup
        roi_sf$date <- time_i
        
        roi_df <- roi_sf %>%
          st_drop_geometry()
        
        #### Export
        saveRDS(roi_df, OUT_FILE)
        
      }
    }
  }
}

#### Append Files
dir.create(file.path(ntl_bm_dir, "aggregated"))

for(unit in c("adm0", 
              "adm1", 
              "adm2",
              "adm3",
              "cities5km",
              "cities10km",
              "citiesbuff")){
  for(time_type in c("annual", "monthly")){
    
    ntl_df <- file.path(ntl_bm_dir, "aggregated_individual_files", unit, time_type) %>%
      list.files(full.names = T,
                 pattern = ".Rds") %>%
      map_df(readRDS)
    
    saveRDS(ntl_df,   file.path(ntl_bm_dir, "aggregated", paste0("ntl_", unit, "_", time_type, ".Rds")))
    write_csv(ntl_df, file.path(ntl_bm_dir, "aggregated", paste0("ntl_", unit, "_", time_type, ".csv")))
    
  }
}

```

# Maps of Nighttime Lights

## Interactive Map

```{r}

#| warning: false
#| message: false

## Load boundaries
adm0_sf <- load_gadm(0)

## Load/prep raster
prep_r <- function(year_i){
  r <- rast(file.path(ntl_bm_dir, "rasters", "annual",
                      paste0("VNP46A4_NearNadir_Composite_Snow_Free_qflag_t",year_i,".tif")))
  r <- r %>% mask(adm0_sf)
  r[][r[] == 0] <- NA
  #r[] <- log(r[] + 1)
  r[] <- log(r[] + 1)
  return(r)
}

r_2012 <- prep_r(2012)
r_2013 <- prep_r(2013)
r_2014 <- prep_r(2014)
r_2015 <- prep_r(2015)
r_2016 <- prep_r(2016)
r_2017 <- prep_r(2017)
r_2018 <- prep_r(2018)
r_2019 <- prep_r(2019)
r_2020 <- prep_r(2020)
r_2021 <- prep_r(2021)
r_2022 <- prep_r(2022)
r_2023 <- prep_r(2023)

## Make map
pal <- colorNumeric(c("yellow", "orange", "red"), unique(c(r_2012[],
                                                           r_2013[],
                                                           r_2014[],
                                                           r_2015[],
                                                           r_2016[],
                                                           r_2017[],
                                                           r_2018[],
                                                           r_2019[],
                                                           r_2020[],
                                                           r_2021[],
                                                           r_2022[],
                                                           r_2023[])),
                    na.color = "transparent")

leaflet() %>%
  addProviderTiles(providers$CartoDB.DarkMatter) %>%
  addRasterImage(r_2012, colors = pal, opacity = 1, group = "2012") %>%
  addRasterImage(r_2013, colors = pal, opacity = 1, group = "2013") %>%
  addRasterImage(r_2014, colors = pal, opacity = 1, group = "2014") %>%
  addRasterImage(r_2015, colors = pal, opacity = 1, group = "2015") %>%
  addRasterImage(r_2016, colors = pal, opacity = 1, group = "2016") %>%
  addRasterImage(r_2017, colors = pal, opacity = 1, group = "2017") %>%
  addRasterImage(r_2018, colors = pal, opacity = 1, group = "2018") %>%
  addRasterImage(r_2019, colors = pal, opacity = 1, group = "2019") %>%
  addRasterImage(r_2020, colors = pal, opacity = 1, group = "2020") %>%
  addRasterImage(r_2021, colors = pal, opacity = 1, group = "2021") %>%
  addRasterImage(r_2022, colors = pal, opacity = 1, group = "2022") %>%
  addRasterImage(r_2023, colors = pal, opacity = 1, group = "2023") %>%
  addLayersControl(
    baseGroups = paste0(2012:2023),
    options = layersControlOptions(collapsed=FALSE)
  )
```


## Static Map of Nighttime Lights

```{r}
#| warning: false
#| message: false

## Load boundaries
adm0_sf <- load_gadm(0)

## Elevation
elev_r <- rast(file.path(data_dir, "elevation", "elev.tiff"))
elev_r <- elev_r %>% crop(adm0_sf) %>% mask(adm0_sf)

## Latest Year
r <- rast(file.path(ntl_bm_dir, "rasters", "annual",
                    paste0("VNP46A4_NearNadir_Composite_Snow_Free_qflag_t",2023,".tif")))
r <- r %>% crop(adm0_sf) %>% mask(adm0_sf)
r[] <- log(r[] + 1)

library(ggnewscale)

r[][r[] <= 1] <- NA
ggplot() +
  geom_spatraster(data = elev_r, show.legend = FALSE) +
  scale_fill_gradient(low = "#49232E",
                      high = "black",
                      na.value = "transparent") +
  ggnewscale::new_scale_fill() +
  geom_spatraster(data = r,
                  aes(fill = t2023)) +
  scale_fill_distiller(palette = "YlGnBl",
                       #midpoint = 2,
                       na.value = "transparent") +
  labs(title = "Nighttime Lights: 2023") +
  coord_sf() +
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        legend.position = "none")
```
# Diagnostics

## Nighttime Lights: Gas Flaring vs Non Gas Flaring

### Trend Figure
```{r}
#| warning: false
#| message: false

ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm0_annual.Rds"))

ntl_df %>%
  dplyr::select(date, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -date) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares",
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares"
  )) %>%
  
  ggplot() +
  geom_col(aes(x = date,
               y = value,
               fill = name),
           stat = "identity", position = 'dodge') +
  labs(fill = NULL,
       x = NULL,
       y = "NTL") +
  scale_fill_manual(values = c(wbg_color_light,
                               wbg_color_dark)) +
  theme_manual
```

### Percent of NTL from Gas Flaring
```{r}
#| warning: false
#| message: false

ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm0_annual.Rds"))

ntl_df %>%
  dplyr::mutate(per_gf = (ntl_gf_10km_sum / ntl_sum) * 100) %>%
  dplyr::select(date, COUNTRY, per_gf) %>%
  pivot_wider(id_cols = date,
              names_from = COUNTRY,
              values_from = per_gf) %>%
  dplyr::rename(Year = date) %>%
  kable()
```

## NTL vs City Population

We use a dataset from [Geonames] that maps the locations of all cities with other 1,000 people. We extract total nighttime lights withing a 5km buffer of each city. The below figure compares nighttime lights with population. 

```{r}
city_df <- readRDS(file.path(data_dir, "cities", "cities_country_df.Rds"))
city_df <- city_df %>%
  dplyr::select(geoname_id, lat, lon)

ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "citiesbuff", "_", "annual", ".Rds")))

ntl_coord_df <- ntl_df %>%
  dplyr::filter(date == 2023) %>%
  left_join(city_df, by = "geoname_id") %>%
  dplyr::mutate(lat = lat %>% as.numeric(),
                lon = lon %>% as.numeric()) %>%
  arrange(ntl_sum)

adm0_sf <- load_gadm(0)

ntl_df %>%
  ggplot() +
  geom_sf(data = adm0_sf) +
  geom_point(data = ntl_coord_df,
             aes(x = lon, 
                 y = lat,
                 fill = log(ntl_sum),
                 size = ntl_sum),
             pch = 21,
             color = "black") +
  labs(size = "Nighttime\nLights",
       title = "Nighttime Lights Across Cities") +
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5)) +
  scale_fill_gradient(low = wbg_color_dark,
                      high = wbg_color_light) +
  guides(fill = "none")

```


```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "cities10km", "_", "annual", ".Rds")))

ntl_df %>%
  dplyr::filter(population > 0,
                date == 2023) %>%
  ggplot(aes(x = log(ntl_sum),
             y = log(population))) +
  geom_point(fill = wbg_color_dark) +
  geom_smooth(method=lm, 
              se=FALSE,
              color = wbg_color_light) +
  stat_cor() +
  theme_minimal() +
  labs(x = "NTL, Logged",
       y = "Population, Logged")
```


## Nighttime Lights as Proxy for GDP
```{r}
# Download GDP
dir.create(file.path(data_dir, "wdi"))
OUT_FILE <- file.path(data_dir, "wdi", "gdp.Rds")

if(!file.exists(OUT_FILE)){
  gdp_df <- WDI(indicator = c("NY.GDP.PCAP.CD",
                              "NY.GDP.MKTP.CD"), 
                country = iso3_code, 
                start = 2012, 
                end = 2023)
  
  saveRDS(gdp_df, OUT_FILE)
} else{
  gdp_df <- readRDS(OUT_FILE)
}

# Correlation with GDP
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm0", "_", "annual", ".Rds")))

ntl_df <- ntl_df %>%
  dplyr::rename(year = date)

# Filter
if(paste(iso3_code, collapse = ";") %in% "SSD;SDN"){
  ntl_df <- ntl_df[ntl_df$COUNTRY %in% "Sudan",]
  gdp_df <- gdp_df[gdp_df$country %in% "Sudan",]
}

ntl_gdp_df <- ntl_df %>%
  left_join(gdp_df, by = "year") %>%
  clean_names()

## Line Plots
ntl_gdp_df %>%
  dplyr::select(year, 
                ntl_sum,
                ntl_gf_10km_sum,
                ntl_nogf_10km_sum,
                ny_gdp_pcap_cd,
                ny_gdp_mktp_cd) %>%
  pivot_longer(cols = -year) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_sum" ~ "Nighttime Lights",
    name == "ntl_gf_10km_sum" ~ "Nighttime Lights:\nGas Flares",
    name == "ntl_nogf_10km_sum" ~ "Nighttime Lights:\nNo Gas Flares",
    name == "ny_gdp_pcap_cd" ~ "GDP, Per Capita",
    name == "ny_gdp_mktp_cd" ~ "GDP"
  )) %>%
  dplyr::mutate(type = name %>% str_detect("GDP")) %>%
  ggplot(aes(x = year,
             y = value,
             color = type)) +
  geom_line(linewidth = 1) +
  facet_wrap(~name,
             scales = "free_y") +
  labs(x = NULL,
       y = NULL) +
  scale_color_manual(values = c(wbg_color_light, wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "none")

## Correlation: GDP
ntl_gdp_df %>%
  dplyr::select(ntl_sum, 
                ntl_gf_10km_sum,
                ntl_nogf_10km_sum,
                ny_gdp_mktp_cd) %>%
  pivot_longer(cols = -c(ny_gdp_mktp_cd)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_sum" ~ "NTL",
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flare",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flare",
  )) %>%
  
  ggplot(aes(x = value,
             y = ny_gdp_mktp_cd)) +
  geom_point() +
  geom_smooth(method=lm, 
              se=FALSE,
              color = wbg_color_light) +
  stat_cor() +
  labs(x = "NTL",
       y = "GDP",
       title = "Correlation of Nighttime Lights with GDP") +
  facet_wrap(~name,
             scales = "free") +
  theme_manual 

## Correlation: GDP
ntl_gdp_df %>%
  dplyr::select(ntl_sum, 
                ntl_gf_10km_sum,
                ntl_nogf_10km_sum,
                ny_gdp_pcap_cd) %>%
  pivot_longer(cols = -c(ny_gdp_pcap_cd)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_sum" ~ "NTL",
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flare",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flare",
  )) %>%
  
  ggplot(aes(x = value,
             y = ny_gdp_pcap_cd)) +
  geom_point() +
  geom_smooth(method=lm, 
              se=FALSE,
              color = wbg_color_light) +
  stat_cor() +
  labs(x = "NTL",
       y = "GDP, per Capita",
       title = "Correlation of Nighttime Lights with per Capita GDP") +
  facet_wrap(~name,
             scales = "free") +
  theme_manual
```


# Trends in Nighttime Lights

## Annual

### Country Level
```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm0", "_", "annual", ".Rds")))

ntl_df %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~COUNTRY) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual
```


```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm0", "_", "annual", ".Rds")))

ntl_df %>%
  dplyr::select(date, COUNTRY, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, COUNTRY)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            linewidth = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~COUNTRY) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm0", "_", "monthly", ".Rds")))

ntl_df %>%
  dplyr::mutate(date = date %>% ymd()) %>%
  dplyr::filter(date >= ymd("2020-01-01")) %>%
  ggplot() +
  geom_vline(xintercept = ymd("2023-04-01"), color = "red") +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~COUNTRY) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual
```

### ADM 1: Top ADM Units
```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm1", "_", "annual", ".Rds")))

top_adm <- ntl_df %>%
  filter(date == 2023) %>%
  arrange(-ntl_sum) %>%
  pull(NAME_1) %>%
  head(12)

ntl_df %>%
  filter(NAME_1 %in% top_adm) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~NAME_1) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual 
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm1", "_", "annual", ".Rds")))

ntl_df %>%
  dplyr::select(date, NAME_1, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, NAME_1)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  filter(NAME_1 %in% top_adm) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            linewidth = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~NAME_1) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm1", "_", "monthly", ".Rds")))

ntl_df %>%
  dplyr::mutate(date = date %>% ymd()) %>%
  dplyr::filter(date >= ymd("2020-01-01")) %>%
  filter(NAME_1 %in% top_adm) %>%
  ggplot() +
  geom_vline(xintercept = ymd("2023-04-01"), color = "red") +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~NAME_1) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual 
```

### ADM 2: Top ADM Units
```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm2", "_", "annual", ".Rds")))

top_adm <- ntl_df %>%
  filter(date == 2023) %>%
  arrange(-ntl_sum) %>%
  pull(NAME_2) %>%
  head(12)

ntl_df %>%
  filter(NAME_2 %in% top_adm) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~NAME_2) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm2", "_", "annual", ".Rds")))

ntl_df %>%
  dplyr::select(date, NAME_2, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, NAME_2)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  filter(NAME_2 %in% top_adm) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            linewidth = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~NAME_2) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm2", "_", "monthly", ".Rds")))

ntl_df %>%
  dplyr::mutate(date = date %>% ymd()) %>%
  dplyr::filter(date >= ymd("2020-01-01")) %>%
  filter(NAME_2 %in% top_adm) %>%
  ggplot() +
  geom_vline(xintercept = ymd("2023-04-01"), color = "red") +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~NAME_2) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual 
```

### Top Cities
```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "citiesbuff", "_", "annual", ".Rds")))

top_adm <- ntl_df %>%
  filter(date == 2023) %>%
  arrange(-ntl_sum) %>%
  pull(geoname_id) %>%
  head(12)

ntl_df %>%
  filter(geoname_id %in% top_adm) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~name) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "cities5km", "_", "annual", ".Rds")))

ntl_df %>%
  dplyr::rename(city_name = name) %>%
  dplyr::select(date, city_name, geoname_id, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, city_name, geoname_id)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  dplyr::filter(geoname_id %in% top_adm) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            size = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~city_name) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "cities5km", "_", "monthly", ".Rds")))

ntl_df %>%
  dplyr::mutate(date = date %>% ymd()) %>%
  dplyr::filter(date >= ymd("2020-01-01")) %>%
  filter(geoname_id %in% top_adm) %>%
  ggplot() +
  geom_vline(xintercept = ymd("2023-04-01"), color = "red") +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~name) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual 
```

# % Change Maps

## ADM 1
```{r}

## Load
adm_sf <- load_gadm(1)
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm1_annual.Rds"))

## Prep data
ntl_wide_df <- ntl_df %>%
  dplyr::filter(date %in% c(pc_base_year:2023)) %>%
  dplyr::select(GID_1, date, ntl_sum) %>%
  dplyr::mutate(date = paste0("yr", date)) %>%
  pivot_wider(names_from = date,
              values_from = ntl_sum,
              id_cols = GID_1)

ntl_wide_df$base_year <- ntl_wide_df[[paste0("yr", pc_base_year)]]

ntl_wide_df <- ntl_wide_df %>%
  dplyr::mutate(pc20 = (yr2020 - base_year)/base_year * 100,
                pc21 = (yr2021 - base_year)/base_year * 100,
                pc22 = (yr2022 - base_year)/base_year * 100,
                pc23 = (yr2023 - base_year)/base_year * 100)

## Map
adm_data_sf <- adm_sf %>%
  left_join(ntl_wide_df, by = "GID_1") %>%
  pivot_longer(cols = c(pc20, pc21, pc22, pc23)) %>%
  dplyr::mutate(name = case_when(
    name == "pc20" ~ paste0("% Change: ",pc_base_year," to 2020"),
    name == "pc21" ~ paste0("% Change: ",pc_base_year," to 2021"),
    name == "pc22" ~ paste0("% Change: ",pc_base_year," to 2022"),
    name == "pc23" ~ paste0("% Change: ",pc_base_year," to 2023")
  ))

adm_data_sf$value[adm_data_sf$value >= 100] <- 100
adm_data_sf$value[adm_data_sf$value <= -100] <- -100

ggplot() +
  geom_sf(data = adm_data_sf,
          aes(fill = value)) +
  facet_wrap(~name) +
  labs(fill = "% Change") +
  scale_fill_gradient2(low = "red",
                       mid = "white",
                       high = "forestgreen",
                       midpoint = 0,
                       limits = c(-100, 100),
                       breaks = c(-100, -50, 0, 50, 100),
                       labels = c("< -100", "-50", "0", "50", "> 100")) +
  theme_manual +
  theme_void() +
  theme(strip.text = element_text(face = "bold"))
```

## ADM 2
```{r}

## Load
adm_sf <- load_gadm(2)
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm2_annual.Rds"))

## Prep data
ntl_wide_df <- ntl_df %>%
  dplyr::filter(date %in% c(pc_base_year:2023)) %>%
  dplyr::select(GID_2, date, ntl_sum) %>%
  dplyr::mutate(date = paste0("yr", date)) %>%
  pivot_wider(names_from = date,
              values_from = ntl_sum,
              id_cols = GID_2)

ntl_wide_df$base_year <- ntl_wide_df[[paste0("yr", pc_base_year)]]

ntl_wide_df <- ntl_wide_df %>%
  dplyr::mutate(pc20 = (yr2020 - base_year)/base_year * 100,
                pc21 = (yr2021 - base_year)/base_year * 100,
                pc22 = (yr2022 - base_year)/base_year * 100,
                pc23 = (yr2023 - base_year)/base_year * 100)

## Map
adm_data_sf <- adm_sf %>%
  left_join(ntl_wide_df, by = "GID_2") %>%
  pivot_longer(cols = c(pc20, pc21, pc22, pc23)) %>%
  dplyr::mutate(name = case_when(
    name == "pc20" ~ paste0("% Change: ",pc_base_year," to 2020"),
    name == "pc21" ~ paste0("% Change: ",pc_base_year," to 2021"),
    name == "pc22" ~ paste0("% Change: ",pc_base_year," to 2022"),
    name == "pc23" ~ paste0("% Change: ",pc_base_year," to 2023")
  ))

adm_data_sf$value[adm_data_sf$value >= 100] <- 100
adm_data_sf$value[adm_data_sf$value <= -100] <- -100

ggplot() +
  geom_sf(data = adm_data_sf,
          aes(fill = value)) +
  facet_wrap(~name) +
  labs(fill = "% Change") +
  scale_fill_gradient2(low = "red",
                       mid = "white",
                       high = "forestgreen",
                       midpoint = 0,
                       limits = c(-100, 100),
                       breaks = c(-100, -50, 0, 50, 100),
                       labels = c("< -100", "-50", "0", "50", "> 100")) +
  theme_manual +
  theme_void() +
  theme(strip.text = element_text(face = "bold"))
```

# Change Before and After Conflict

## % Change ADM2 Map

Percent change in NTL: Jan - March 2023, vs May - Dec 2023

```{r}

## Load
adm_sf <- load_gadm(2)
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm2_monthly.Rds"))

## Prep data
ntl_wide_df <- ntl_df %>%
  dplyr::mutate(date = date %>% ymd(),
                year = date %>% year(),
                date_month = date %>% month(),
                period = case_when(
                  date_month %in% 1:3 ~ "ntl_before",
                  date_month %in% 5:12 ~ "ntl_after"
                )) %>%
  dplyr::filter(year == 2023,
                !is.na(period)) %>%
  group_by(GID_2, period) %>%
  dplyr::summarise(ntl_sum = mean(ntl_sum)) %>%
  ungroup() %>%
  pivot_wider(id_cols = GID_2,
              names_from = period,
              values_from = ntl_sum) %>%
  dplyr::mutate(pc = (ntl_after - ntl_before) / ntl_before * 100)

adm_sf <- adm_sf %>%
  left_join(ntl_wide_df, by = "GID_2")

adm_sf$value[adm_sf$pc >= 100] <- 100
adm_sf$value[adm_sf$pc <= -100] <- -100

ggplot() +
  geom_sf(data = adm_sf,
          aes(fill = pc)) +
  labs(fill = "% Change") +
  scale_fill_gradient2(low = "red",
                       mid = "white",
                       high = "forestgreen",
                       midpoint = 0,
                       limits = c(-100, 100),
                       breaks = c(-100, -50, 0, 50, 100),
                       labels = c("< -100", "-50", "0", "50", "> 100")) +
  theme_manual +
  theme_void() +
  theme(strip.text = element_text(face = "bold"))
```

## % Change ADM3 Map

Percent change in NTL: Jan - March 2023, vs May - Dec 2023

```{r}

## Load
adm_sf <- load_gadm(3)
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm3_monthly.Rds"))

## Prep data
ntl_wide_df <- ntl_df %>%
  dplyr::mutate(date = date %>% ymd(),
                year = date %>% year(),
                date_month = date %>% month(),
                period = case_when(
                  date_month %in% 1:3 ~ "ntl_before",
                  date_month %in% 5:12 ~ "ntl_after"
                )) %>%
  dplyr::filter(year == 2023,
                !is.na(period)) %>%
  group_by(GID_3, period) %>%
  dplyr::summarise(ntl_sum = mean(ntl_sum)) %>%
  ungroup() %>%
  pivot_wider(id_cols = GID_3,
              names_from = period,
              values_from = ntl_sum) %>%
  dplyr::mutate(pc = (ntl_after - ntl_before) / ntl_before * 100)

adm_sf <- adm_sf %>%
  left_join(ntl_wide_df, by = "GID_3")

adm_sf$value[adm_sf$pc >= 100] <- 100
adm_sf$value[adm_sf$pc <= -100] <- -100

ggplot() +
  geom_sf(data = adm_sf,
          aes(fill = pc)) +
  labs(fill = "% Change") +
  scale_fill_gradient2(low = "red",
                       mid = "white",
                       high = "forestgreen",
                       midpoint = 0,
                       limits = c(-100, 100),
                       breaks = c(-100, -50, 0, 50, 100),
                       labels = c("< -100", "-50", "0", "50", "> 100")) +
  theme_manual +
  theme_void() +
  theme(strip.text = element_text(face = "bold"))
```

## % Change Pixel Map

```{r}
adm0_sf <- load_gadm(0)

r_before <- file.path(ntl_bm_dir, "rasters", "monthly") %>%
  list.files(full.names = T) %>%
  str_subset("2023_01|2023_02|2023_03") %>%
  rast() %>%
  mean()

r_after <- file.path(ntl_bm_dir, "rasters", "monthly") %>%
  list.files(full.names = T) %>%
  str_subset("2023_05|2023_06|2023_07|2023_08|2023_09|2023_10|2023_11|2023_12") %>%
  rast() %>%
  mean()

r_c <- (r_after - r_before) / r_before * 100
r_c[r_before < 1] <- NA
r_c[r_c >= 100]  <- 100
r_c[r_c <= -100] <- -100
r_c <- r_c %>% crop(adm0_sf) %>% mask(adm0_sf)

# ggplot() +
#   geom_spatraster(data = r_c) +
#   geom_sf(data = adm0_sf,
#           fill = NA,
#           color = "black") +
#   scale_fill_gradient2(low = "red", mid = "white", high = "green", 
#                        midpoint = 0,
#                        na.value = "white") +
#   theme_void()

library(leaflet)
library(leaflet.extras)

# Define color palette for raster values
pal <- colorNumeric(palette = colorRamp(c("red", "black", "green")),
                    domain = values(r_c), na.color = NA)

# Create leaflet map
leaflet() %>%
  addTiles() %>%
  addRasterImage(r_c, colors = pal, opacity = 1) %>%
  #addPolygons(data = adm0_sf, color = "black", fill = NA) %>%
  addLegend(pal = pal, values = values(r_c), title = "% Change") %>%
  setView(lng = 0, lat = 0, zoom = 2)  # Adjust longitude, latitude, and zoom as needed

```

## Table of % Change

The below tables show the percent change in lights after the conflict, quarterly, relative to 2022.

```{r}
#### Prep Rasters
r_before <- file.path(ntl_bm_dir, "rasters", "annual") %>%
  list.files(full.names = T) %>%
  str_subset("2022") %>%
  rast() 

r_2023_04_06 <- file.path(ntl_bm_dir, "rasters", "monthly") %>%
  list.files(full.names = T) %>%
  str_subset("2023_04|2023_05|2023_06") %>%
  rast() %>%
  mean()

r_2023_07_09 <- file.path(ntl_bm_dir, "rasters", "monthly") %>%
  list.files(full.names = T) %>%
  str_subset("2023_07|2023_08|2023_09") %>%
  rast() %>%
  mean()

r_2023_10_12 <- file.path(ntl_bm_dir, "rasters", "monthly") %>%
  list.files(full.names = T) %>%
  str_subset("2023_10|2023_11|2023_12") %>%
  rast() %>%
  mean()

r_2024_01_03 <- file.path(ntl_bm_dir, "rasters", "monthly") %>%
  list.files(full.names = T) %>%
  str_subset("2023_01|2023_02|2023_03") %>%
  rast() %>%
  mean()

r_2024_04_06 <- file.path(ntl_bm_dir, "rasters", "monthly") %>%
  list.files(full.names = T) %>%
  str_subset("2023_04|2023_05|2023_06") %>%
  rast() %>%
  mean()

make_table <- function(roi_sf, unit_name){
  
  roi_sf$ntl_2022       <- exact_extract(r_before,     roi_sf, "sum", progress = F)
  roi_sf$ntl_2023_04_06 <- exact_extract(r_2023_04_06, roi_sf, "sum", progress = F)
  roi_sf$ntl_2023_07_09 <- exact_extract(r_2023_07_09, roi_sf, "sum", progress = F)
  roi_sf$ntl_2023_10_12 <- exact_extract(r_2023_10_12, roi_sf, "sum", progress = F)
  roi_sf$ntl_2024_01_03 <- exact_extract(r_2024_01_03, roi_sf, "sum", progress = F)
  roi_sf$ntl_2024_04_06 <- exact_extract(r_2024_04_06, roi_sf, "sum", progress = F)
  
  roi_sf <- roi_sf %>%
    dplyr::mutate(pc_2023_04_06 = (ntl_2023_04_06 - ntl_2022) / ntl_2022 * 100,
                  pc_2023_07_09 = (ntl_2023_07_09 - ntl_2022) / ntl_2022 * 100,
                  pc_2023_10_12 = (ntl_2023_10_12 - ntl_2022) / ntl_2022 * 100,
                  pc_2024_01_03 = (ntl_2024_01_03 - ntl_2022) / ntl_2022 * 100,
                  pc_2024_04_06 = (ntl_2024_04_06 - ntl_2022) / ntl_2022 * 100) %>%
    st_drop_geometry() %>%
    dplyr::filter(ntl_2022 > 0)
  
  roi_sf$Name <- roi_sf[[unit_name]]
  
  roi_sf %>%
    dplyr::select(Name, ntl_2022, contains("pc_")) %>%
    dplyr::mutate(across(c(ntl_2022, contains("pc_")), ~ round(.x, 2))) %>%
    dplyr::rename("NTL: 2022" = ntl_2022,
                  "% Change: 2023 Q2"= pc_2023_04_06,
                  "% Change: 2023 Q3"= pc_2023_07_09,
                  "% Change: 2023 Q4"= pc_2023_10_12,
                  "% Change: 2024 Q1"= pc_2024_01_03,
                  "% Change: 2024 Q2"= pc_2024_04_06) %>%
    DT::datatable() 
}
```

### Country Level

```{r}
roi_sf <- load_gadm(0)
make_table(roi_sf, "COUNTRY")
```

### ADM 1

```{r}
roi_sf <- load_gadm(1)
make_table(roi_sf, "NAME_1")
```

### ADM 2

```{r}
roi_sf <- load_gadm(2)
make_table(roi_sf, "NAME_2")
```

### ADM 3

```{r}
roi_sf <- load_gadm(3)
make_table(roi_sf, "NAME_3")
```

### Cities

```{r}
roi_sf <- readRDS(file.path(data_dir, "cities", "cities_country_buff_sf.Rds"))
make_table(roi_sf, "ascii_name")
```

# Conflict vs NTL: ADM2

We calculate the percent change in NTL and compare the percent change in more and less conflict affected locations (eg, ADM2). Percent change in NTL from Jan - March 2023 to May - Dec 2023.

```{r}
## Load data
gadm2_sf <- load_gadm(2)

nt1_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm1_monthly.Rds"))
nt2_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm2_monthly.Rds"))

acled_df <- read_csv(file.path(data_dir, "acled",
                               "raw_sudan_acled_20230101_20241101.csv"))

## Prep conflict
acled_sf <- acled_df %>%
  dplyr::filter(event_date >= ymd("2023-04-15")) %>%
  st_as_sf(coords = c("longitude", "latitude"),
           crs = 4326) %>%
  dplyr::select(fatalities) %>%
  st_join(gadm2_sf)

acled_1_df <- acled_sf %>%
  st_drop_geometry() %>%
  group_by(GID_1) %>%
  dplyr::summarise(n_events = n(),
                   fatalities = sum(fatalities))

acled_2_df <- acled_sf %>%
  st_drop_geometry() %>%
  group_by(GID_2) %>%
  dplyr::summarise(n_events = n(),
                   fatalities = sum(fatalities))

## Change in NTL
nt1_change_df <- nt1_df %>%
  dplyr::filter(date >= ymd("2023-01-01"),
                date != ymd("2023-04-01"),
                date <= ymd("2023-12-01")) %>%
  dplyr::mutate(post = as.numeric(date >= ymd("2023-04-01"))) %>%
  group_by(GID_1, post) %>%
  dplyr::summarise(ntl_sum = mean(ntl_sum)) %>%
  ungroup() %>%
  pivot_wider(id_cols = GID_1,
              names_from = post,
              values_from = ntl_sum) %>%
  dplyr::mutate(pc = (`1` - `0`) / `0` ) %>%
  left_join(acled_1_df, by = "GID_1")

nt2_change_df <- nt2_df %>%
  dplyr::filter(date >= ymd("2023-01-01"),
                date != ymd("2023-04-01"),
                date <= ymd("2023-12-01")) %>%
  dplyr::mutate(post = as.numeric(date >= ymd("2023-04-01"))) %>%
  group_by(GID_2, post) %>%
  dplyr::summarise(ntl_sum = mean(ntl_sum)) %>%
  ungroup() %>%
  pivot_wider(id_cols = GID_2,
              names_from = post,
              values_from = ntl_sum) %>%
  dplyr::mutate(pc = (`1` - `0`) / `0` * 100 ) %>%
  left_join(acled_2_df, by = "GID_2")

nt2_change_df <- nt2_change_df %>%
  dplyr::mutate(conflict_zone = as.numeric(fatalities >= 10),
                conflict_zone_str = ifelse(conflict_zone == 1,
                                           "Conflict\n(>= 10 fatalities)",
                                           "No/Less Conflict\n(< 10 fatalities)"))
nt2_change_df$pc[nt2_change_df$pc >= 100] <- 100
nt2_change_df$pc[nt2_change_df$pc <= -100] <- -100

nt2_change_df %>%
  dplyr::filter(!is.na(conflict_zone)) %>%
  ggplot() +
  geom_boxplot(aes(x = pc,
                   y = conflict_zone_str),
               fill = "darkorange") +
    geom_vline(xintercept = 0,
             color = "red") +
  labs(x = "Percent Change in NTL",
       y = NULL) + 
  theme_classic2()
```


