---
title: "Nighttime Lights Analysis"
author: "Data Lab"
date: "9/20/2024"
format:
  html:
    toc: true
    code-fold: true
knitr:
  opts_chunk:
    warning: false
    message: false
---

```{r}
#| code-fold: true

library(tidyverse)
library(lubridate)

# REQUIRED PARAMETERS ----------------------------------------------------------
proj_dir <- "~/Desktop/NTL Analysis" # Directory for data; code will create this folder
iso3_code    <- "NGA"          # ISO3 code for country
nasa_bearer <- "BEARER-TOKEN"  # NASA Bearer Token

# OPTIONAL PARAMETERS ----------------------------------------------------------

## Years to Query
year_start <- 2012
year_end <- Sys.Date() %>% year()

## Months to query
month_start <- "2012-01-01"    # Start month to query 
month_end   <- Sys.Date() %>%  # End month to query
  floor_date(unit = "months") %>% 
  as.character()  

# Base year for % change maps
pc_base_year <- 2019 
```

Nighttime lights have become a commonly used resource to estimate changes in local economic activity. This document analyzes nighttime lights within the country.

# Data

We use nighttime lights data from [VIIRS Black Marble](https://blackmarble.gsfc.nasa.gov/). Raw nighttime lights data requires correction due to cloud cover and stray light, such as lunar light. The Black Marble dataset applies [advanced algorithms](https://www.sciencedirect.com/science/article/pii/S003442571830110X) to correct raw nighttime light values and calibrate data so that trends in lights over time can be meaningfully analyzed. From VIIRS Black Marble, we use data from January 2012 through present—where data is available at a 500-meter resolution.

# Methodology

Within different units of analysis (e.g, administrative units) we use the sum of nighttime lights. Where relevant, we distinguish between lights generated from gas flaring and lights removing gas flaring. We use the [World Bank's Global Gas Flaring Tracker](https://www.worldbank.org/en/programs/gasflaringreduction/global-flaring-data) which indicates the location of gas flaring locations. When removing gas flaring lights, we remove lights within 10km of a gas flaring location; when looking at lights in gas flaring locations, we take the sum of lights within 10km of gas flaring locations. 

# Code Setup

```{r}
#| warning: false
#| message: false

#### Packages
library(tidyverse)
library(sf)
library(leaflet)
library(leaflet.providers)
library(ggpubr)
library(terra)
library(tidyterra)
library(gtools)
library(readxl)
library(janitor)
library(geodata)
library(exactextractr)
library(blackmarbler)
library(dplyr)
library(readxl)
library(janitor)
library(ggpubr)
library(WDI)
library(kableExtra)
library(terra)
library(DT)
library(countrycode)
library(haven)

## Iso2 Code
iso2_code <- countrycode(iso3_code, origin = "iso3c", destination = "iso2c")

## Filepath
data_dir       <- file.path(proj_dir, "data")
ntl_bm_dir     <- file.path(data_dir, "Nighttime Lights BlackMarble")
gadm_dir       <- file.path(data_dir, "GADM")
gas_flare_dir  <- file.path(data_dir, "gas-flaring")

## Create
dir.create(proj_dir)
dir.create(data_dir)
dir.create(ntl_bm_dir)
dir.create(gadm_dir)
dir.create(gas_flare_dir)

#### Bearer
nasa_bearer <- read_csv("~/Dropbox/bearer_bm.csv") %>%
  pull(token)

#### Theme
theme_manual <- theme_classic2() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        plot.title = element_text(face = "bold"))

#### GADM Boundaries
for(i in 0:3){
  for(iso3_i in iso3_code){
    gadm(iso3_code, level=i, path = file.path(gadm_dir))
  }
} 

gadm_n_files <- file.path(gadm_dir, 'gadm') %>%
  list.files(pattern = "*.rds") %>%
  length()

if(gadm_n_files == 1) gadm_names <- c("adm0")
if(gadm_n_files == 2) gadm_names <- c("adm0", "adm1")
if(gadm_n_files == 3) gadm_names <- c("adm0", "adm1", "adm2")
if(gadm_n_files == 4) gadm_names <- c("adm0", "adm1", "adm2", "adm3")

#### Load GADM Functions
load_gadm <- function(i){
  roi_sf <- file.path(gadm_dir, 'gadm') %>%
    list.files(full.names = T) %>%
    str_subset(paste0("_",i,"_pk.rds")) %>%
    map_df(function(file_path){
      readRDS(file_path) %>%
        st_as_sf()
    })
}

#### Elevation
dir.create(file.path(data_dir, "elevation"))

OUT_PATH <- file.path(data_dir, "elevation", "elev.tiff")

if(!file.exists(OUT_PATH)){
  
  if(length(iso3_code) == 1){
    elev_r <- elevation_30s(country=iso3_code, path=tempdir())
    
  } else{
    elev_list <- lapply(iso3_code, function(iso3_code_i){
      elevation_30s(country=iso3_code_i, path=tempdir())
    }) 
    elev_r <- do.call(mosaic, c(elev_list, fun="mean"))
  }
  
  writeRaster(elev_r, OUT_PATH)
  
}

#### Months to Query
ntl_months <- seq.Date(from = ymd(month_start), 
                       to = ymd(month_end), 
                       by = "month") %>%
  as.character()

#### Colors
wbg_color_light <- "#1AA1DB"
wbg_color_dark <- "#02224B"
```

# Prep Data

## Download Cities

We use a dataset from [Geonames](https://public.opendatasoft.com/explore/dataset/geonames-all-cities-with-a-population-1000/export/?flg=en-us&disjunctive.cou_name_en&sort=name) which indicates the location of cities with over 1,000 people. 

```{r}
#| warning: false
#| message: false

dir.create(file.path(data_dir, "cities"))

## Cities
OUT_FILE_ALL <- file.path(data_dir, "cities", "cities_all.Rds")

if(!file.exists(OUT_FILE_ALL)){
  
  # Set the URL and destination path
  url <- "https://jeodpp.jrc.ec.europa.eu/ftp/jrc-opendata/GHSL/GHS_STAT_UCDB2015MT_GLOBE_R2019A/V1-2/GHS_STAT_UCDB2015MT_GLOBE_R2019A_V1_2.zip"
  destfile <- file.path(data_dir, "cities", "GHS_STAT_UCDB2015MT_GLOBE_R2019A_V1_2.zip")
  
  # Download the zip file
  download.file(url, destfile, mode = "wb")
  
  # Unzip the downloaded file
  unzip(destfile, exdir = file.path(data_dir, "cities", "GHS_STAT_data"))
  
  # Delete unneeded files
  unlink(destfile)
  unlink(file.path(data_dir, "cities", "GHS_STAT_data", "GHS_STAT_UCDB2015MT_GLOBE_R2019A", "GHS_STAT_UCDB2015MT_GLOBE_R2019A_V1_0_web.pdf"))
  unlink(file.path(data_dir, "cities", "GHS_STAT_data", "GHS_STAT_UCDB2015MT_GLOBE_R2019A", "GHS_STAT_UCDB2015MT_GLOBE_R2019A_V1_2_short.csv"))
  unlink(file.path(data_dir, "cities", "GHS_STAT_data", "GHS_STAT_UCDB2015MT_GLOBE_R2019A", "GHS_STAT_UCDB2015MT_GLOBE_R2019A_V1_2.csv"))
  unlink(file.path(data_dir, "cities", "GHS_STAT_data", "GHS_STAT_UCDB2015MT_GLOBE_R2019A", "GHS_STAT_UCDB2015MT_GLOBE_R2019A_V1_2.xls"))
  
  ## Load data
  city_sf <- read_sf(file.path(data_dir, "cities", "GHS_STAT_data", "GHS_STAT_UCDB2015MT_GLOBE_R2019A",
                               "GHS_STAT_UCDB2015MT_GLOBE_R2019A_V1_2.gpkg"))
  
  #city_p_sf <- read_sf(file.path(data_dir, "cities", "GHS_STAT_data", "GHS_STAT_UCDB2015MT_GLOBE_R2019A",
  #                               "GHS_STAT_UCDB2015MT_GLOBE_R2019A_V1_2_short_pnt.gpkg"))
  
  ## Restrict to country
  city_sf <- city_sf[city_sf$CTR_MN_ISO %in% iso3_code,]
  
  city_sf <- city_sf %>%
    dplyr::select(ID_HDC_G0, CTR_MN_ISO, UC_NM_MN, P15)
  
  saveRDS(city_sf, OUT_FILE_ALL)
}

```


## Download and Append gas flaring

We download and append gas flaring data from the [World Bank Global Gas Flaring Database](https://datacatalog.worldbank.org/search/dataset/0037743).

```{r}
#| warning: false
#| message: false

dir.create(file.path(gas_flare_dir, "rawdata"))
dir.create(file.path(gas_flare_dir, "finaldata"))

#### Download data
# https://datacatalog.worldbank.org/search/dataset/0037743
if(!file.exists(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2017.xlsx"))){
  download.file(url = "https://datacatalogfiles.worldbank.org/ddh-published/0037743/DR0045623/viirs_global_flaring_d.7_slope_0.029353_2017_web_v1.xlsx?versionId=2023-01-18T20:03:32.2273754Z", 
                destfile = file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2017.xlsx"), 
                mode = "wb")
}

if(!file.exists(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2018.xlsx"))){
  download.file(url = "https://datacatalogfiles.worldbank.org/ddh-published/0037743/DR0045622/viirs_global_flaring_d.7_slope_0.029353_2018_web.xlsx?versionId=2023-01-18T20:02:43.3965005Z", 
                destfile = file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2018.xlsx"), 
                mode = "wb")
}

if(!file.exists(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2019.xlsx"))){
  download.file(url = "https://datacatalogfiles.worldbank.org/ddh-published/0037743/DR0045621/viirs_global_flaring_d.7_slope_0.029353_2019_web_v20201114-3.xlsx?versionId=2023-01-18T20:03:09.2456111Z", 
                destfile = file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2019.xlsx"), 
                mode = "wb")
}

if(!file.exists(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2020.xlsx"))){
  download.file(url = "https://datacatalogfiles.worldbank.org/ddh-published/0037743/DR0084248/2020%20Global%20Gas%20Flaring%20Volumes.xlsx?versionId=2023-01-18T20:03:53.8309309Z", 
                destfile = file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2020.xlsx"), 
                mode = "wb")
}

if(!file.exists(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2021.xlsx"))){
  download.file(url = "https://datacatalogfiles.worldbank.org/ddh-published/0037743/DR0087112/2021%20Global%20Gas%20Flaring%20Volumes.xlsx?versionId=2023-01-18T20:02:21.4951166Z", 
                destfile = file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2021.xlsx"), 
                mode = "wb")
}

#### Append data
clean_data <- function(x){
  x %>% 
    clean_names() %>% 
    dplyr::filter(iso_code %in% iso3_code)
} 

df_2021 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2021.xlsx"), 2) %>% clean_data()

df_2020_1 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2020.xlsx"), 1) %>% clean_data()
df_2020_2 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2020.xlsx"), 2) %>% clean_data()
df_2020_3 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2020.xlsx"), 3) %>% clean_data()

df_2019 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2019.xlsx"), 1) %>% clean_data()

df_2018_4 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2018.xlsx"), 4) %>% clean_data()
df_2018_5 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2018.xlsx"), 5) %>% clean_data()
df_2018_6 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2018.xlsx"), 6) %>% clean_data()

df_2017_1 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2017.xlsx"), 1) %>% clean_data()
df_2017_2 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2017.xlsx"), 2) %>% clean_data()
df_2017_3 <- read_xlsx(file.path(gas_flare_dir, "rawdata", "viirs_global_flaring_2017.xlsx"), 3) %>% clean_data()

gs_df <- bind_rows(
  df_2021,
  df_2020_1,
  df_2020_2,
  df_2020_3,
  df_2019,
  df_2018_4,
  df_2018_5,
  df_2018_6,
  df_2017_1,
  df_2017_2,
  df_2017_3
)

if(nrow(gs_df) > 0){
  gs_df <- gs_df %>%
    dplyr::select(latitude, longitude) %>%
    distinct() %>%
    dplyr::mutate(uid = 1:n())
  
  gs_sf <- gs_df %>% 
    st_as_sf(coords = c("longitude", "latitude"),
             crs = 4326)
  
  gs_5km_sf <- gs_sf %>%
    st_buffer(dist = 5000)
  
  gs_10km_sf <- gs_sf %>%
    st_buffer(dist = 10000)
  
} else{
  
  gs_df <- data.frame(NULL)
  gs_5km_sf <- data.frame(NULL)
  gs_10km_sf <- data.frame(NULL)
  
}

saveRDS(gs_df, file.path(gas_flare_dir, "finaldata", "gas_flare_locations.Rds"))
saveRDS(gs_5km_sf, file.path(gas_flare_dir, "finaldata", "gas_flare_locations_5km.Rds"))
saveRDS(gs_10km_sf, file.path(gas_flare_dir, "finaldata", "gas_flare_locations_10km.Rds"))
```

## Download Nighttime Lights

```{r}
#| warning: false
#| message: false

#### NTL
roi_sf <- load_gadm(0)

dir.create(file.path(ntl_bm_dir, "rasters"))
dir.create(file.path(ntl_bm_dir, "rasters", "annual"))
dir.create(file.path(ntl_bm_dir, "rasters", "monthly"))
dir.create(file.path(ntl_bm_dir, "rasters", "daily"))

r <- bm_raster(roi_sf = roi_sf,
               product_id = "VNP46A4",
               date = year_start:year_end,
               bearer = nasa_bearer,
               output_location_type = "file",
               file_dir = file.path(ntl_bm_dir, "rasters", "annual"))

r <- bm_raster(roi_sf = roi_sf,
               product_id = "VNP46A3",
               date = ntl_months,
               bearer = nasa_bearer,
               output_location_type = "file",
               file_dir = file.path(ntl_bm_dir, "rasters", "monthly"))
```

## Aggregate Nighttime Lights
```{r}
#| warning: false
#| message: false

# Individual Files -------------------------------------------------------------
dir.create(file.path(ntl_bm_dir, "aggregated_individual_files"))

#### Setup loops
for(unit in c(gadm_names,
              "cities")){
  for(time_type in c("annual", "monthly")){
    
    if(time_type == "annual"){
      time_vec <- year_start:year_end
    }
    
    if(time_type == "monthly"){
      time_vec <- ntl_months
    }
    
    for(time_i in time_vec){
      
      #### Only process if file doesn't exist
      dir.create(file.path(ntl_bm_dir, "aggregated_individual_files", unit))
      dir.create(file.path(ntl_bm_dir, "aggregated_individual_files", unit, time_type))
      
      OUT_FILE <- file.path(ntl_bm_dir, "aggregated_individual_files", unit, time_type, 
                            paste0("ntl_", time_i, ".Rds"))
      
      if(!file.exists(OUT_FILE)){
        
        #### Load Unit
        if(unit == "adm0"){
          roi_sf <- load_gadm(0)
        }
        
        if(unit == "adm1"){
          roi_sf <- load_gadm(1)
        }
        
        if(unit == "adm2"){
          roi_sf <- load_gadm(2)
        }
        
        if(unit == "adm3"){
          roi_sf <- load_gadm(3)
        }
        
        if(unit == "cities"){
          roi_sf <- readRDS(file.path(data_dir, "cities", "cities_all.Rds"))
        }
        
        #### Load Raster
        if(time_type == "annual"){
          bm_product <- "VNP46A4"
          time_raster_i <- time_i 
        }
        
        if(time_type == "monthly"){
          bm_product <- "VNP46A3"
          
          time_raster_i <- time_i %>% 
            substring(1, 7) %>%
            str_replace_all("-", "_")
        }
        
        r_path <- file.path(ntl_bm_dir, "rasters", time_type, 
                            paste0(bm_product,"_NearNadir_Composite_Snow_Free_qflag_t",time_raster_i,".tif"))
        
        if(file.exists(r_path)){
          
          ntl_r <- rast(r_path)
          
          #### Gas Flaring Mask
          gs_5km_sf  <- readRDS(file.path(gas_flare_dir, "finaldata", "gas_flare_locations_5km.Rds"))
          gs_10km_sf <- readRDS(file.path(gas_flare_dir, "finaldata", "gas_flare_locations_10km.Rds"))
          
          if(nrow(gs_5km_sf) > 0){
            ntl_gf_5km_r   <- mask(ntl_r, gs_5km_sf)
            ntl_nogf_5km_r <- mask(ntl_r, gs_5km_sf, inverse = T)
            
            ntl_gf_10km_r   <- mask(ntl_r, gs_10km_sf)
            ntl_nogf_10km_r <- mask(ntl_r, gs_10km_sf, inverse = T)
          }
          
          #### Extract NTL
          roi_sf$ntl_sum          <- exact_extract(ntl_r, roi_sf, "sum", progress = F)
          
          if(nrow(gs_5km_sf) > 0){
            roi_sf$ntl_gf_5km_sum   <- exact_extract(ntl_gf_5km_r, roi_sf, "sum", progress = F)
            roi_sf$ntl_nogf_5km_sum <- exact_extract(ntl_nogf_5km_r, roi_sf, "sum", progress = F)
            
            roi_sf$ntl_gf_10km_sum   <- exact_extract(ntl_gf_10km_r, roi_sf, "sum", progress = F)
            roi_sf$ntl_nogf_10km_sum <- exact_extract(ntl_nogf_10km_r, roi_sf, "sum", progress = F)
          } else{
            roi_sf$ntl_gf_5km_sum   <- 0
            roi_sf$ntl_nogf_5km_sum <- roi_sf$ntl_sum
            
            roi_sf$ntl_gf_10km_sum   <- 0
            roi_sf$ntl_nogf_10km_sum <- roi_sf$ntl_sum
          }
          
          #### Cleanup
          roi_sf$date <- time_i
          
          roi_df <- roi_sf %>%
            st_drop_geometry()
          
          #### Export
          saveRDS(roi_df, OUT_FILE)
          
        }
      }
    }
  }
}

#### Append Files
dir.create(file.path(ntl_bm_dir, "aggregated"))

for(unit in c(gadm_names,
              "cities")){
  for(time_type in c("annual", "monthly")){
    
    ntl_df <- file.path(ntl_bm_dir, "aggregated_individual_files", unit, time_type) %>%
      list.files(full.names = T,
                 pattern = ".Rds") %>%
      map_df(readRDS)
    
    saveRDS(ntl_df,   file.path(ntl_bm_dir, "aggregated", paste0("ntl_", unit, "_", time_type, ".Rds")))
    write_csv(ntl_df, file.path(ntl_bm_dir, "aggregated", paste0("ntl_", unit, "_", time_type, ".csv")))
    write_dta(ntl_df, file.path(ntl_bm_dir, "aggregated", paste0("ntl_", unit, "_", time_type, ".dta")))
    
  }
}

```

# Maps of Nighttime Lights

## Interactive Map

```{r}

#| warning: false
#| message: false

## Load boundaries
adm0_sf <- load_gadm(0)

## Load/prep raster
prep_r <- function(year_i){
  r <- rast(file.path(ntl_bm_dir, "rasters", "annual",
                      paste0("VNP46A4_NearNadir_Composite_Snow_Free_qflag_t",year_i,".tif")))
  r <- r %>% mask(adm0_sf)
  r[][r[] == 0] <- NA
  #r[] <- log(r[] + 1)
  r[] <- log(r[] + 1)
  return(r)
}

r_2012 <- prep_r(2012)
r_2013 <- prep_r(2013)
r_2014 <- prep_r(2014)
r_2015 <- prep_r(2015)
r_2016 <- prep_r(2016)
r_2017 <- prep_r(2017)
r_2018 <- prep_r(2018)
r_2019 <- prep_r(2019)
r_2020 <- prep_r(2020)
r_2021 <- prep_r(2021)
r_2022 <- prep_r(2022)
r_2023 <- prep_r(2023)

## Make map
pal <- colorNumeric(c("yellow", "orange", "red"), unique(c(r_2012[],
                                                           r_2013[],
                                                           r_2014[],
                                                           r_2015[],
                                                           r_2016[],
                                                           r_2017[],
                                                           r_2018[],
                                                           r_2019[],
                                                           r_2020[],
                                                           r_2021[],
                                                           r_2022[],
                                                           r_2023[])),
                    na.color = "transparent")

leaflet() %>%
  addProviderTiles(providers$CartoDB.DarkMatter) %>%
  addRasterImage(r_2012, colors = pal, opacity = 1, group = "2012") %>%
  addRasterImage(r_2013, colors = pal, opacity = 1, group = "2013") %>%
  addRasterImage(r_2014, colors = pal, opacity = 1, group = "2014") %>%
  addRasterImage(r_2015, colors = pal, opacity = 1, group = "2015") %>%
  addRasterImage(r_2016, colors = pal, opacity = 1, group = "2016") %>%
  addRasterImage(r_2017, colors = pal, opacity = 1, group = "2017") %>%
  addRasterImage(r_2018, colors = pal, opacity = 1, group = "2018") %>%
  addRasterImage(r_2019, colors = pal, opacity = 1, group = "2019") %>%
  addRasterImage(r_2020, colors = pal, opacity = 1, group = "2020") %>%
  addRasterImage(r_2021, colors = pal, opacity = 1, group = "2021") %>%
  addRasterImage(r_2022, colors = pal, opacity = 1, group = "2022") %>%
  addRasterImage(r_2023, colors = pal, opacity = 1, group = "2023") %>%
  addLayersControl(
    baseGroups = paste0(2012:2023),
    options = layersControlOptions(collapsed=FALSE)
  )
```


## Static Map of Nighttime Lights

```{r}
#| warning: false
#| message: false

## Load boundaries
adm0_sf <- load_gadm(0)

## Elevation
elev_r <- rast(file.path(data_dir, "elevation", "elev.tiff"))
elev_r <- elev_r %>% crop(adm0_sf) %>% mask(adm0_sf)

## Latest Year
r <- rast(file.path(ntl_bm_dir, "rasters", "annual",
                    paste0("VNP46A4_NearNadir_Composite_Snow_Free_qflag_t",2023,".tif")))
r <- r %>% crop(adm0_sf) %>% mask(adm0_sf)
r[] <- log(r[] + 1)

library(ggnewscale)

r[][r[] <= 1] <- NA
ggplot() +
  geom_spatraster(data = elev_r, show.legend = FALSE) +
  scale_fill_gradient(low = "#49232E",
                      high = "black",
                      na.value = "transparent") +
  ggnewscale::new_scale_fill() +
  geom_spatraster(data = r,
                  aes(fill = t2023)) +
  scale_fill_distiller(palette = "YlGnBl",
                       #midpoint = 2,
                       na.value = "transparent") +
  labs(title = "Nighttime Lights: 2023") +
  coord_sf() +
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        legend.position = "none")
```
# Diagnostics

## Nighttime Lights: Gas Flaring vs Non Gas Flaring

### Trend Figure
```{r}
#| warning: false
#| message: false

ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm0_annual.Rds"))

if("ntl_nogf_10km_sum" %in% names(ntl_df)){
  
  ntl_df %>%
    dplyr::select(date, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
    pivot_longer(cols = -date) %>%
    dplyr::mutate(name = case_when(
      name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares",
      name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares"
    )) %>%
    
    ggplot() +
    geom_col(aes(x = date,
                 y = value,
                 fill = name),
             stat = "identity", position = 'dodge') +
    labs(fill = NULL,
         x = NULL,
         y = "NTL") +
    scale_fill_manual(values = c(wbg_color_light,
                                 wbg_color_dark)) +
    theme_manual
} else{
  cat("No gas flaring in the country")
}
```

### Percent of NTL from Gas Flaring
```{r}
#| warning: false
#| message: false

if("ntl_nogf_10km_sum" %in% names(ntl_df)){
  ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm0_annual.Rds"))
  
  ntl_df %>%
    dplyr::mutate(per_gf = (ntl_gf_10km_sum / ntl_sum) * 100) %>%
    dplyr::select(date, COUNTRY, per_gf) %>%
    pivot_wider(id_cols = date,
                names_from = COUNTRY,
                values_from = per_gf) %>%
    dplyr::rename(Year = date) %>%
    kable()
  
}
```

## NTL vs City Population

We use a dataset from [GHSL](https://data.jrc.ec.europa.eu/dataset/53473144-b88c-44bc-b4a3-4583ed1f547e) that maps the locations of all cities. We extract total nighttime lights each city. The below figure compares nighttime lights with population. 

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "cities", "_", "annual", ".Rds")))


ntl_df %>%
  dplyr::filter(P15 > 0,
                date == 2015) %>%
  ggplot(aes(x = log(ntl_sum),
             y = log(P15))) +
  geom_point(fill = wbg_color_dark) +
  geom_smooth(method=lm, 
              se=FALSE,
              color = wbg_color_light) +
  stat_cor() +
  theme_minimal() +
  labs(x = "NTL, Logged",
       y = "Population, Logged")
```


## Nighttime Lights as Proxy for GDP
```{r}
# Download GDP
dir.create(file.path(data_dir, "wdi"))
OUT_FILE <- file.path(data_dir, "wdi", "gdp.Rds")

if(!file.exists(OUT_FILE)){
  gdp_df <- WDI(indicator = c("NY.GDP.PCAP.CD",
                              "NY.GDP.MKTP.CD"), 
                country = iso3_code, 
                start = 2012, 
                end = 2023)
  
  saveRDS(gdp_df, OUT_FILE)
} else{
  gdp_df <- readRDS(OUT_FILE)
}

# Correlation with GDP
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "adm0", "_", "annual", ".Rds")))

ntl_df <- ntl_df %>%
  dplyr::rename(year = date)

# Filter
if(paste(iso3_code, collapse = ";") %in% "SSD;SDN"){
  ntl_df <- ntl_df[ntl_df$COUNTRY %in% "Sudan",]
  gdp_df <- gdp_df[gdp_df$country %in% "Sudan",]
}

ntl_gdp_df <- ntl_df %>%
  left_join(gdp_df, by = "year") %>%
  clean_names()

## Line Plots
ntl_gdp_df[,names(ntl_gdp_df) %in% 
             c("year", 
               "ntl_sum",
               "ntl_gf_10km_sum",
               "ntl_nogf_10km_sum",
               #"ny_gdp_pcap_cd",
               "ny_gdp_mktp_cd")] %>%
  pivot_longer(cols = -year) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_sum" ~ "Nighttime Lights",
    name == "ntl_gf_10km_sum" ~ "Nighttime Lights:\nGas Flares",
    name == "ntl_nogf_10km_sum" ~ "Nighttime Lights:\nNo Gas Flares",
    name == "ny_gdp_pcap_cd" ~ "GDP, Per Capita",
    name == "ny_gdp_mktp_cd" ~ "GDP"
  )) %>%
  dplyr::mutate(type = name %>% str_detect("GDP")) %>%
  ggplot(aes(x = year,
             y = value,
             color = type)) +
  geom_line(linewidth = 1) +
  facet_wrap(~name,
             scales = "free_y") +
  labs(x = NULL,
       y = NULL) +
  scale_color_manual(values = c(wbg_color_light, wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "none")

## Correlation: GDP
ntl_gdp_df[,names(ntl_gdp_df) %in% 
             c("ntl_sum", 
               "ntl_gf_10km_sum",
               "ntl_nogf_10km_sum",
               "ny_gdp_mktp_cd")] %>%
  pivot_longer(cols = -c(ny_gdp_mktp_cd)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_sum" ~ "NTL",
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flare",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flare",
  )) %>%
  
  ggplot(aes(x = value,
             y = ny_gdp_mktp_cd)) +
  geom_point() +
  geom_smooth(method=lm, 
              se=FALSE,
              color = wbg_color_light) +
  stat_cor() +
  labs(x = "NTL",
       y = "GDP",
       title = "Correlation of Nighttime Lights with GDP") +
  facet_wrap(~name,
             scales = "free") +
  theme_manual 

## Correlation: GDP
# ntl_gdp_df %>%
#   dplyr::select(ntl_sum, 
#                 ntl_gf_10km_sum,
#                 ntl_nogf_10km_sum,
#                 ny_gdp_pcap_cd) %>%
#   pivot_longer(cols = -c(ny_gdp_pcap_cd)) %>%
#   dplyr::mutate(name = case_when(
#     name == "ntl_sum" ~ "NTL",
#     name == "ntl_gf_10km_sum" ~ "NTL: Gas Flare",
#     name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flare",
#   )) %>%
#   
#   ggplot(aes(x = value,
#              y = ny_gdp_pcap_cd)) +
#   geom_point() +
#   geom_smooth(method=lm, 
#               se=FALSE,
#               color = wbg_color_light) +
#   stat_cor() +
#   labs(x = "NTL",
#        y = "GDP, per Capita",
#        title = "Correlation of Nighttime Lights with per Capita GDP") +
#   facet_wrap(~name,
#              scales = "free") +
#   theme_manual
```


# Trends in Nighttime Lights

## Country Level
```{r}
if("adm0" %in% gadm_names){
  
  ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                              paste0("ntl_", "adm0", "_", "annual", ".Rds")))
  
  ntl_df %>%
    ggplot() +
    geom_col(aes(x = date, y = ntl_sum),
             fill = wbg_color_light) +
    facet_wrap(~COUNTRY) +
    labs(x = NULL,
         y = "Nighttime Lights") +
    theme_manual
  
} else{
  cat("Country level polygon does not exist in GADM")
}
```

```{r}
if("adm0" %in% gadm_names){
  ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                              paste0("ntl_", "adm0", "_", "annual", ".Rds")))
  
  ntl_df %>%
    dplyr::select(date, COUNTRY, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
    pivot_longer(cols = -c(date, COUNTRY)) %>%
    dplyr::mutate(name = case_when(
      name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
      name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
    )) %>%
    ggplot() +
    geom_line(aes(x = date,
                  y = value,
                  color = name),
              linewidth = 1) +
    labs(color = NULL,
         x = NULL,
         y = "NTL") +
    facet_wrap(~COUNTRY) +
    scale_color_manual(values = c(wbg_color_light,
                                  wbg_color_dark)) +
    theme_manual +
    theme(legend.position = "top")
}
```

```{r}
if("adm0" %in% gadm_names){
  ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                              paste0("ntl_", "adm0", "_", "monthly", ".Rds")))
  
  ntl_df %>%
    dplyr::mutate(date = date %>% ymd()) %>%
    ggplot() +
    geom_col(aes(x = date, y = ntl_sum),
             fill = wbg_color_light) +
    facet_wrap(~COUNTRY) +
    labs(x = NULL,
         y = "Nighttime Lights") +
    theme_manual
}
```

## ADM 1: Top ADM Units
```{r}
if("adm1" %in% gadm_names){
  ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                              paste0("ntl_", "adm1", "_", "annual", ".Rds")))
  
  top_adm <- ntl_df %>%
    dplyr::filter(date == max(date)) %>%
    arrange(-ntl_sum) %>%
    pull(NAME_1) %>%
    head(12)
  
  ntl_df %>%
    filter(NAME_1 %in% top_adm) %>%
    ggplot() +
    geom_col(aes(x = date, y = ntl_sum),
             fill = wbg_color_light) +
    facet_wrap(~NAME_1) +
    labs(x = NULL,
         y = "Nighttime Lights") +
    theme_manual 
} else{
  cat("ADM1 GADM file does not exist")
}
```

```{r}
if("adm1" %in% gadm_names){
  ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                              paste0("ntl_", "adm1", "_", "annual", ".Rds")))
  
  ntl_df %>%
    dplyr::select(date, NAME_1, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
    pivot_longer(cols = -c(date, NAME_1)) %>%
    dplyr::mutate(name = case_when(
      name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
      name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
    )) %>%
    filter(NAME_1 %in% top_adm) %>%
    ggplot() +
    geom_line(aes(x = date,
                  y = value,
                  color = name),
              linewidth = 1) +
    labs(color = NULL,
         x = NULL,
         y = "NTL") +
    facet_wrap(~NAME_1) +
    scale_color_manual(values = c(wbg_color_light,
                                  wbg_color_dark)) +
    theme_manual +
    theme(legend.position = "top")
}
```

```{r}
if("adm1" %in% gadm_names){
  ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                              paste0("ntl_", "adm1", "_", "monthly", ".Rds")))
  
  ntl_df %>%
    dplyr::mutate(date = date %>% ymd()) %>%
    filter(NAME_1 %in% top_adm) %>%
    ggplot() +
    geom_col(aes(x = date, y = ntl_sum),
             fill = wbg_color_light) +
    facet_wrap(~NAME_1) +
    labs(x = NULL,
         y = "Nighttime Lights") +
    theme_manual 
}
```

## ADM 2: Top ADM Units
```{r}
if("adm2" %in% gadm_names){
  ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                              paste0("ntl_", "adm2", "_", "annual", ".Rds")))
  
  top_adm <- ntl_df %>%
    dplyr::filter(date == max(date)) %>%
    arrange(-ntl_sum) %>%
    pull(NAME_2) %>%
    head(12)
  
  ntl_df %>%
    filter(NAME_2 %in% top_adm) %>%
    ggplot() +
    geom_col(aes(x = date, y = ntl_sum),
             fill = wbg_color_light) +
    facet_wrap(~NAME_2) +
    labs(x = NULL,
         y = "Nighttime Lights") +
    theme_manual
} else{
  cat("ADM2 GADM polygon does not exist")
}
```

```{r}
if("adm2" %in% gadm_names){
  ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                              paste0("ntl_", "adm2", "_", "annual", ".Rds")))
  
  ntl_df %>%
    dplyr::select(date, NAME_2, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
    pivot_longer(cols = -c(date, NAME_2)) %>%
    dplyr::mutate(name = case_when(
      name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
      name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
    )) %>%
    filter(NAME_2 %in% top_adm) %>%
    ggplot() +
    geom_line(aes(x = date,
                  y = value,
                  color = name),
              linewidth = 1) +
    labs(color = NULL,
         x = NULL,
         y = "NTL") +
    facet_wrap(~NAME_2) +
    scale_color_manual(values = c(wbg_color_light,
                                  wbg_color_dark)) +
    theme_manual +
    theme(legend.position = "top")
}
```

```{r}
if("adm2" %in% gadm_names){
  ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                              paste0("ntl_", "adm2", "_", "monthly", ".Rds")))
  
  ntl_df %>%
    dplyr::filter(date >= ymd("2020-01-01")) %>%
    dplyr::mutate(date = ymd(date)) %>%
    filter(NAME_2 %in% top_adm) %>%
    ggplot() +
    geom_col(aes(x = date, y = ntl_sum),
             fill = wbg_color_light) +
    facet_wrap(~NAME_2) +
    labs(x = NULL,
         y = "Nighttime Lights") +
    theme_manual 
}
```

## Top Cities
```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "cities", "_", "annual", ".Rds")))

top_adm <- ntl_df %>%
  dplyr::filter(date == max(date)) %>%
  arrange(-ntl_sum) %>%
  pull(ID_HDC_G0) %>%
  head(12)

ntl_df %>%
  filter(ID_HDC_G0 %in% top_adm) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~UC_NM_MN) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "cities", "_", "annual", ".Rds")))

ntl_df %>%
  dplyr::rename(city_name = UC_NM_MN) %>%
  dplyr::select(date, city_name, ID_HDC_G0, ntl_gf_10km_sum, ntl_nogf_10km_sum) %>%
  pivot_longer(cols = -c(date, city_name, ID_HDC_G0)) %>%
  dplyr::mutate(name = case_when(
    name == "ntl_gf_10km_sum" ~ "NTL: Gas Flares",
    name == "ntl_nogf_10km_sum" ~ "NTL: No Gas Flares"
  )) %>%
  dplyr::filter(ID_HDC_G0 %in% top_adm) %>%
  ggplot() +
  geom_line(aes(x = date,
                y = value,
                color = name),
            size = 1) +
  labs(color = NULL,
       x = NULL,
       y = "NTL") +
  facet_wrap(~city_name) +
  scale_color_manual(values = c(wbg_color_light,
                                wbg_color_dark)) +
  theme_manual +
  theme(legend.position = "top")
```

```{r}
ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", 
                            paste0("ntl_", "cities", "_", "monthly", ".Rds")))

ntl_df %>%
  dplyr::mutate(date = date %>% ymd()) %>%
  filter(ID_HDC_G0 %in% top_adm) %>%
  ggplot() +
  geom_col(aes(x = date, y = ntl_sum),
           fill = wbg_color_light) +
  facet_wrap(~UC_NM_MN) +
  labs(x = NULL,
       y = "Nighttime Lights") +
  theme_manual 
```

# % Change Maps

## ADM 1
```{r}

if("adm1" %in% gadm_names){
  
  ## Load
  adm_sf <- load_gadm(1)
  ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm1_annual.Rds"))
  
  ## Prep data
  ntl_wide_df <- ntl_df %>%
    dplyr::filter(date %in% c(pc_base_year:2023)) %>%
    dplyr::select(GID_1, date, ntl_sum) %>%
    dplyr::mutate(date = paste0("yr", date)) %>%
    pivot_wider(names_from = date,
                values_from = ntl_sum,
                id_cols = GID_1)
  
  ntl_wide_df$base_year <- ntl_wide_df[[paste0("yr", pc_base_year)]]
  
  ntl_wide_df <- ntl_wide_df %>%
    dplyr::mutate(pc20 = (yr2020 - base_year)/base_year * 100,
                  pc21 = (yr2021 - base_year)/base_year * 100,
                  pc22 = (yr2022 - base_year)/base_year * 100,
                  pc23 = (yr2023 - base_year)/base_year * 100)
  
  ## Map
  adm_data_sf <- adm_sf %>%
    left_join(ntl_wide_df, by = "GID_1") %>%
    pivot_longer(cols = c(pc20, pc21, pc22, pc23)) %>%
    dplyr::mutate(name = case_when(
      name == "pc20" ~ paste0("% Change: ",pc_base_year," to 2020"),
      name == "pc21" ~ paste0("% Change: ",pc_base_year," to 2021"),
      name == "pc22" ~ paste0("% Change: ",pc_base_year," to 2022"),
      name == "pc23" ~ paste0("% Change: ",pc_base_year," to 2023")
    ))
  
  adm_data_sf$value[adm_data_sf$value >= 100] <- 100
  adm_data_sf$value[adm_data_sf$value <= -100] <- -100
  
  ggplot() +
    geom_sf(data = adm_data_sf,
            aes(fill = value)) +
    facet_wrap(~name) +
    labs(fill = "% Change") +
    scale_fill_gradient2(low = "red",
                         mid = "white",
                         high = "forestgreen",
                         midpoint = 0,
                         limits = c(-100, 100),
                         breaks = c(-100, -50, 0, 50, 100),
                         labels = c("< -100", "-50", "0", "50", "> 100")) +
    theme_manual +
    theme_void() +
    theme(strip.text = element_text(face = "bold"))
  
} else{
  cat("ADM1 GADM polygon does not exist")
}
```

## ADM 2
```{r}
if("adm2" %in% gadm_names){
  ## Load
  adm_sf <- load_gadm(2)
  ntl_df <- readRDS(file.path(ntl_bm_dir, "aggregated", "ntl_adm2_annual.Rds"))
  
  ## Prep data
  ntl_wide_df <- ntl_df %>%
    dplyr::filter(date %in% c(pc_base_year:2023)) %>%
    dplyr::select(GID_2, date, ntl_sum) %>%
    dplyr::mutate(date = paste0("yr", date)) %>%
    pivot_wider(names_from = date,
                values_from = ntl_sum,
                id_cols = GID_2)
  
  ntl_wide_df$base_year <- ntl_wide_df[[paste0("yr", pc_base_year)]]
  
  ntl_wide_df <- ntl_wide_df %>%
    dplyr::mutate(pc20 = (yr2020 - base_year)/base_year * 100,
                  pc21 = (yr2021 - base_year)/base_year * 100,
                  pc22 = (yr2022 - base_year)/base_year * 100,
                  pc23 = (yr2023 - base_year)/base_year * 100)
  
  ## Map
  adm_data_sf <- adm_sf %>%
    left_join(ntl_wide_df, by = "GID_2") %>%
    pivot_longer(cols = c(pc20, pc21, pc22, pc23)) %>%
    dplyr::mutate(name = case_when(
      name == "pc20" ~ paste0("% Change: ",pc_base_year," to 2020"),
      name == "pc21" ~ paste0("% Change: ",pc_base_year," to 2021"),
      name == "pc22" ~ paste0("% Change: ",pc_base_year," to 2022"),
      name == "pc23" ~ paste0("% Change: ",pc_base_year," to 2023")
    ))
  
  adm_data_sf$value[adm_data_sf$value >= 100] <- 100
  adm_data_sf$value[adm_data_sf$value <= -100] <- -100
  
  ggplot() +
    geom_sf(data = adm_data_sf,
            aes(fill = value)) +
    facet_wrap(~name) +
    labs(fill = "% Change") +
    scale_fill_gradient2(low = "red",
                         mid = "white",
                         high = "forestgreen",
                         midpoint = 0,
                         limits = c(-100, 100),
                         breaks = c(-100, -50, 0, 50, 100),
                         labels = c("< -100", "-50", "0", "50", "> 100")) +
    theme_manual +
    theme_void() +
    theme(strip.text = element_text(face = "bold"))
} else{
  cat("ADM2 GADM polygon does not exist")
}
```

